---
title: "Simpsons Guest Stars"
output: github_document
---

# Load Data and Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse) #data management and ggplot
library(igraph) # bipartite networks
library(RColorBrewer)

simpsons <- readr::read_delim("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-27/simpsons-guests.csv", delim = "|", quote = "")

simpsons
```

# Inspect data

* *variable*	class	description
* *season*	integer	Season of the show
* *number*	character	Episode number
* *production_code*	character	Production code for the episode
* *episode_title*	character	Episode Title
* *guest_star*	character	Guest star (actual name)
* *role*	character	Role in the show, either a character or themself

```{r}
#how many roles have each guest star played?
simpsons %>%
  group_by(guest_star) %>%
  summarize(length(role))

# How many guest stars per season
simpsons %>%
  group_by(as.numeric(season)) %>% #make season numeric so in proper numerical order
  tally()
```
# Grab data for SNA

*only picking last 5 seasons for ease of viewing connections*
```{r}
#select data to graph
guests<- simpsons %>%
  filter(season %in% 26:30) %>%
  select(guest_star, number)

#create contigency matrix
guest_mat<- as.data.frame.matrix(table(guests))

#graph incidence
guests_inc1<- graph.incidence(guest_mat)

# creat bipartite networks
guests_pr1<- bipartite.projection(guests_inc1)

#select just project1 from bipartite (vertex names= guest star names)
collab_guests<- guests_pr1$proj1

#crude plot, will need some serious formatting
plot(collab_guests)

#filter for degree of less than 5
less_deg<- which(degree(collab_guests)<5)

#pull out those who do not have many connections
collab_guests_deg5<- delete.vertices(collab_guests, less_deg)

#formatting for better visual
plot(collab_guests_deg5, 
     vertex.label=V(collab_guests_deg5)$name, 
     vertex.size=degree(collab_guests_deg5)/3, 
     vertex.color="darkred", 
     layout=layout_with_kk, 
     main="Simpsons Guest Starring Network (Seasons 26 through 30)", 
     vertex.label.cex = .60)

#look at total degree by guest
 prom.df <- data.frame(
            names= V(collab_guests_deg5)$name,
            deg.guests = degree(collab_guests_deg5))
 
 
prom.df<- prom.df %>%
   filter(deg.guests>8)
   
#bar charts_ top degree
p <- ggplot(prom.df, aes(x = reorder(names, +deg.guests), y = deg.guests)) +
         geom_bar(stat = "identity", fill="Salmon") + theme_minimal()
p+ coord_flip() + ggtitle("Prominent Guest Stars in the Simpsons' Guest Star Network\n (Seasons 26-30)") + labs(x="Guest Star Names", y= "Number of Episodes shared with other Guests (Degree)")

```

